---
title: "Free Charging Trial Survey"
author: "Michael Maness"
date: "November 18, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(apollo)
library(ggplot2)
```

## Load Dataset


```{r load-data}
data.wide <- read.csv('trial_data_freecharge_coded.csv')
data.wide$ID <- 1:nrow(data.wide)
design <- read.csv('experimental_design_trial.csv')
```

## Table Manipulation

Let me changed the data into a long format based on scenario

```{r data-long}
data <- data.wide %>% pivot_longer(starts_with('Scenario'),
                      names_to = c('ScenarioNum'),
                      names_pattern = 'Scenario(.*)',
                      values_to = 'Choice')
```

Now I want to get rid of the rows from the unseen scenarios.

```{r data-elim-scenarios}
data <- data %>% filter(!is.na(Choice))
```

Lets combine the scenario design data with the choice data:

```{r combine-data-design}
colnames(design)[1] <- 'ScenarioNum'
data$ScenarioNum <- as.integer(data$ScenarioNum)
data <- left_join(data,design, by='ScenarioNum')
write.csv("data")
```

write.csv("data")
## Descriptive Statistics

Lets get some stats on scenario choice

```{r stats-01}
scenario_counts <- data %>% group_by(ScenarioNum) %>% count(Choice)
as.data.frame(scenario_counts %>% pivot_wider(names_from = 'Choice', values_from = 'n'))

ggplot(data, aes(x=Choice,color="green")) + geom_bar(fill = "darkblue") + facet_wrap(~ScenarioNum)
scenario_counts

```


## Apollo Setup

```{r apollo-setup}
### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName  ="Charger_Choice_MNL_linear_price",
  modelDescr ="Simple MNL model with linear price",
  indivID    ="ID"
)
```

## Apollo: Analysis of Choice

```{r apollo-choice-analysis}
ABC_avail = rep.int(x = 1, times = nrow(data))
database <- as.data.frame(data)

choiceAnalysis_settings <- list(
  alternatives = c(A=1, B=2, C=3),
  avail        = list(A=ABC_avail, B=ABC_avail, C=ABC_avail),
  choiceVar    = database$Choice,
  explanators  = database[,c("Gender")]
)

apollo_choiceAnalysis(choiceAnalysis_settings, apollo_control, database)
```

## Apollo: Define Model Parameters

```{r model-params}
### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(asc_A   = 0,
              asc_B   = 0,
              asc_C   = 0,
              beta_price = 0,
              beta_free = 0,
              beta_chargetime = 0,
              beta_detourtime = 0,
              beta_restaurant = 0,
              beta_convstore = 0,
              beta_mall = 0,
              beta_interactionalt1=0,
              beta_interactionalt2=0,
              beta_interactionalt3=0)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_A")

apollo_inputs = apollo_validateInputs()
```

## Apollo: Likelihood Function and Model Definition

```{r apollo-likelihood}
apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  V[['A']]  = asc_A + beta_price * alt1.price + 
                      beta_free * (alt1.price == 0) +
                      beta_chargetime * alt1.charge_time + 
                      beta_detourtime * alt1.detour_time +
                      beta_convstore * (alt1.amenities == -1) +
    beta_interactionalt1*((alt1.price == 0)*(alt1.amenities == -1))+ 
                      beta_restaurant * (alt1.amenities == 1) +beta_interactionalt2*((alt1.price == 0) * (alt1.amenities == 1))+
                      beta_mall * (alt1.amenities == 3)+beta_interactionalt3*((alt1.price == 0) * (alt1.amenities == 3))
  V[['B']]  = asc_B + beta_price * alt2.price + 
                      beta_free * (alt2.price == 0) +
                      beta_chargetime * alt2.charge_time + 
                     beta_detourtime * alt2.detour_time +
                      beta_convstore * (alt2.amenities == -1) + beta_interactionalt1*((alt2.price == 0)*(alt2.amenities == -1))+
                      beta_restaurant * (alt2.amenities == 1) + beta_interactionalt2*((alt2.price == 0) * (alt2.amenities == 1))+
                      beta_mall * (alt2.amenities == 3)+beta_interactionalt3*((alt2.price == 0) * (alt2.amenities == 3))
  V[['C']]  = asc_C + beta_price * alt3.price + 
                      beta_free * (alt3.price == 0) +
                      beta_chargetime * alt3.charge_time + 
                      beta_detourtime * alt3.detour_time +
                      beta_convstore * (alt3.amenities == -1) + beta_interactionalt1*((alt3.price == 0)*(alt3.amenities == -1))+ 
                      beta_restaurant * (alt3.amenities == 1) + beta_interactionalt2*((alt3.price == 0) * (alt3.amenities == 1))+ 
                      beta_mall * (alt3.amenities == 3)+beta_interactionalt3*((alt3.price == 0) * (alt3.amenities == 3))
  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives  = c(A=1, B=2, C=3), 
    avail         = list(A=ABC_avail, B=ABC_avail, C=ABC_avail), 
    choiceVar     = Choice,
    V             = V
  )
  
  ### Compute probabilities using MNL model
  P[['model']] = apollo_mnl(mnl_settings, functionality)

  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)

  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}
```

## Apollo: Model Estimation and Output

```{r apollo-model-estimation}
model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

apollo_modelOutput(model)
```

## Second estimation with quadratic price

Now I want to try the estimation again with a quadratic price relationship instead

```{r quadratic-price-estimation}
### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName  ="Charger_Choice_MNL_quadratic_price",
  modelDescr ="Simple MNL model with linear price",
  indivID    ="ID"
)

ABC_avail = rep.int(x = 1, times = nrow(data))
database <- as.data.frame(data)

choiceAnalysis_settings <- list(
  alternatives = c(A=1, B=2, C=3),
  avail        = list(A=ABC_avail, B=ABC_avail, C=ABC_avail),
  choiceVar    = database$Choice,
  explanators  = database[,c("Gender")]
)

apollo_choiceAnalysis(choiceAnalysis_settings, apollo_control, database)

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(asc_A   = 0,
              asc_B   = 0,
              asc_C   = 0,
              beta_price = 0,
              beta_price_squared = 0,
              beta_free = 0,
              beta_chargetime = 0,
              beta_detourtime = 0,
              beta_restaurant = 0,
              beta_convstore = 0,
              beta_mall = 0)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_A")

apollo_inputs = apollo_validateInputs()

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  V[['A']]  = asc_A + beta_price * alt1.price + 
                      beta_price_squared * alt1.price^2 +
                      beta_free * (alt1.price == 0) +
                      beta_chargetime * alt1.charge_time + 
                      beta_detourtime * alt1.detour_time +
                      beta_convstore * (alt1.amenities == -1) + 
                      beta_restaurant * (alt1.amenities == 1) + 
                      beta_mall * (alt1.amenities == 3)
  V[['B']]  = asc_B + beta_price * alt2.price + 
                      beta_price_squared * alt2.price^2 +
                      beta_free * (alt2.price == 0) +
                      beta_chargetime * alt2.charge_time + 
                      beta_detourtime * alt2.detour_time +
                      beta_convstore * (alt2.amenities == -1) + 
                      beta_restaurant * (alt2.amenities == 1) + 
                      beta_mall * (alt2.amenities == 3)
  V[['C']]  = asc_C + beta_price * alt3.price + 
                      beta_price_squared * alt3.price^2 +
                      beta_free * (alt3.price == 0) +
                      beta_chargetime * alt3.charge_time + 
                      beta_detourtime * alt3.detour_time +
                      beta_convstore * (alt3.amenities == -1) + 
                      beta_restaurant * (alt3.amenities == 1) + 
                      beta_mall * (alt3.amenities == 3)
  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives  = c(A=1, B=2, C=3), 
    avail         = list(A=ABC_avail, B=ABC_avail, C=ABC_avail), 
    choiceVar     = Choice,
    V             = V
  )
  
  ### Compute probabilities using MNL model
  P[['model']] = apollo_mnl(mnl_settings, functionality)

  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)

  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

model2 = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

apollo_modelOutput(model2)
```


## Third estimation with quadratic price and no ASCs

```{r estimation-quadratic-noASC}
### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName  ="Charger_Choice_MNL_quadratic_price_noASC",
  modelDescr ="Simple MNL model with linear price",
  indivID    ="ID"
)

ABC_avail = rep.int(x = 1, times = nrow(data))
database <- as.data.frame(data)

choiceAnalysis_settings <- list(
  alternatives = c(A=1, B=2, C=3),
  avail        = list(A=ABC_avail, B=ABC_avail, C=ABC_avail),
  choiceVar    = database$Choice,
  explanators  = database[,c("Gender")]
)

apollo_choiceAnalysis(choiceAnalysis_settings, apollo_control, database)

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(asc_A   = 0,
              asc_B   = 0,
              asc_C   = 0,
              beta_price = 0,
              beta_price_squared = 0,
              beta_free = 0,
              beta_chargetime = 0,
              beta_detourtime = 0,
              beta_restaurant = 0,
              beta_convstore = 0,
              beta_mall = 0)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_A", "asc_B", "asc_C")

apollo_inputs = apollo_validateInputs()

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  V[['A']]  = asc_A + beta_price * alt1.price + 
                      beta_price_squared * alt1.price^2 +
                      beta_free * (alt1.price == 0) +
                      beta_chargetime * alt1.charge_time + 
                      beta_detourtime * alt1.detour_time +
                      beta_convstore * (alt1.amenities == -1) + 
                      beta_restaurant * (alt1.amenities == 1) + 
                      beta_mall * (alt1.amenities == 3)
  V[['B']]  = asc_B + beta_price * alt2.price + 
                      beta_price_squared * alt2.price^2 +
                      beta_free * (alt2.price == 0) +
                      beta_chargetime * alt2.charge_time + 
                      beta_detourtime * alt2.detour_time +
                      beta_convstore * (alt2.amenities == -1) + 
                      beta_restaurant * (alt2.amenities == 1) + 
                      beta_mall * (alt2.amenities == 3)
  V[['C']]  = asc_C + beta_price * alt3.price + 
                      beta_price_squared * alt3.price^2 +
                      beta_free * (alt3.price == 0) +
                      beta_chargetime * alt3.charge_time + 
                      beta_detourtime * alt3.detour_time +
                      beta_convstore * (alt3.amenities == -1) + 
                      beta_restaurant * (alt3.amenities == 1) + 
                      beta_mall * (alt3.amenities == 3)
  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives  = c(A=1, B=2, C=3), 
    avail         = list(A=ABC_avail, B=ABC_avail, C=ABC_avail), 
    choiceVar     = Choice,
    V             = V
  )
  
  ### Compute probabilities using MNL model
  P[['model']] = apollo_mnl(mnl_settings, functionality)

  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)

  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

model3 = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

apollo_modelOutput(model3)
```

## Fourth estimation with Linear Price and No ASC

```{r estimation-linear-noASC}
### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName  ="Charger_Choice_MNL_linear_price_noASC",
  modelDescr ="Simple MNL model with linear price",
  indivID    ="ID"
)

ABC_avail = rep.int(x = 1, times = nrow(data))
database <- as.data.frame(data)

choiceAnalysis_settings <- list(
  alternatives = c(A=1, B=2, C=3),
  avail        = list(A=ABC_avail, B=ABC_avail, C=ABC_avail),
  choiceVar    = database$Choice,
  explanators  = database[,c("Gender")]
)

apollo_choiceAnalysis(choiceAnalysis_settings, apollo_control, database)

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(asc_A   = 0,
              asc_B   = 0,
              asc_C   = 0,
              beta_price = 0,
              beta_free = 0,
              beta_chargetime = 0,
              beta_detourtime = 0,
              beta_restaurant = 0,
              beta_convstore = 0,
              beta_mall = 0)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_A", "asc_B", "asc_C")

apollo_inputs = apollo_validateInputs()

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  V[['A']]  = asc_A + beta_price * alt1.price + 
                      beta_free * (alt1.price == 0) +
                      beta_chargetime * alt1.charge_time + 
                      beta_detourtime * alt1.detour_time +
                      beta_convstore * (alt1.amenities == -1) + 
                      beta_restaurant * (alt1.amenities == 1) + 
                      beta_mall * (alt1.amenities == 3)
  V[['B']]  = asc_B + beta_price * alt2.price + 
                      beta_free * (alt2.price == 0) +
                      beta_chargetime * alt2.charge_time + 
                      beta_detourtime * alt2.detour_time +
                      beta_convstore * (alt2.amenities == -1) + 
                      beta_restaurant * (alt2.amenities == 1) + 
                      beta_mall * (alt2.amenities == 3)
  V[['C']]  = asc_C + beta_price * alt3.price + 
                      beta_free * (alt3.price == 0) +
                      beta_chargetime * alt3.charge_time + 
                      beta_detourtime * alt3.detour_time +
                      beta_convstore * (alt3.amenities == -1) + 
                      beta_restaurant * (alt3.amenities == 1) + 
                      beta_mall * (alt3.amenities == 3)
  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives  = c(A=1, B=2, C=3), 
    avail         = list(A=ABC_avail, B=ABC_avail, C=ABC_avail), 
    choiceVar     = Choice,
    V             = V
  )
  
  ### Compute probabilities using MNL model
  P[['model']] = apollo_mnl(mnl_settings, functionality)

  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)

  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

model4 = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

apollo_modelOutput(model4)
```

